\vspace{-1em}Guidance, please! Towards a framework for RDF-based constraint languages.

### 2015-07-07 paper/paper 2/dc-2015.tex
section 2.1

In 2013, the \emph{W3C} organized the \emph{RDF Validation Workshop}\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}
where experts from industry, government, and academia discussed first RDF validation use cases. 
%for RDF data validation and constraint formulation.
In 2014, two working groups on RDF validation have been established: 
the \emph{W3C RDF Data Shapes Working Group}\footnote{\url{http://www.w3.org/2014/rds/charter}} and the \emph{DCMI RDF Application Profiles Task Group}\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}. 
We collected the findings of these working groups and initiated a database of RDF validation requirements\footnote{Online available at \url{http://purl.org/net/rdf-validation}}
with the intention to collaboratively collect case studies, use cases, requirements, and solutions in a comprehensive and structured way \cite{BoschEckert2014}. 
Based on our work in the \emph{DCMI} and in cooperation with the \emph{W3C} working group,
we published by today 81 requirements to validate RDF data and to formulate constraints; 
%each of them corresponding to a constraint type which in most cases can be expressed by multiple constraint languages. 
each of them corresponds to a constraint type from which concrete constraints are instantiated to be checked on RDF data. 
%and which in most cases can be expressed by multiple constraint languages.
%Requirements/constraint types are uniquely identified by alphanumeric technical identifiers.
%\emph{Minimum qualified cardinality restrictions} which corresponds to the requirement \emph{R-75}, e.g., is a constraint type which corresponds to the requirement {\small\emph{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}}.
We recently published a technical report (serving as first appendix of this paper) in which we explain each constraint type in detail and give examples for each represented by different constraint languages \cite{BoschNolleAcarEckert2015}.
More than 10 constraint languages with different syntax and semantics exist or are currently developed. 
We focus on the five most promising ones on being the standard like
\emph{Shape Expressions (ShEx)}, \emph{Resource Shapes (ReSh)}, and \emph{Description Set Profiles (DSP)}. 
The \emph{Web Ontology Language} \emph{(OWL)} in its current version \emph{OWL 2} can also be used as a constraint language.
With its direct support of validation via \emph{SPARQL}, the \emph{SPARQL Inferencing Notation (SPIN)} is very popular and certainly plays an important role for the future development in this field. 
%\emph{SPIN} is particularly interesting as a means to validate arbitrary constraint languages by mapping them to \emph{SPARQL} \cite{BoschEckert2014-2}.

@@

\begin{table}[H]
	\centering
	  %\caption{Constraint Type Specific Expressivity}
		\caption{Constraint Type Specific Expressivity of Constraint Languages}
	  \scriptsize
		\begin{tabular}{c|c|c|c|c}
      %\textbf{Expressivity Class} & $Expr(\overline{\mathcal{CT}_R})$ (42) & $Expr(\mathcal{CT}_R)$ (32) \\		
			\emph{DSP} & \emph{ReSh} & \emph{ShEx} & \emph{OWL 2} & \emph{SPIN} \\	
      \hline
			17.3 (14) & 25.9 (21) & 29.6 (24) & 67.9 (55) & \textbf{100.0} \textbf{(81)} 
		\end{tabular}
	\label{tab:constraint-type-specific-expressivity}
\end{table}

The fact that \emph{SPIN} supports all and \emph{OWL 2} covers 2/3 of all constraint types
emphasizes the significant role \emph{SPIN} and \emph{OWL 2} play for the future development of constraint languages.
\emph{SPARQL} and therefore \emph{SPIN}, the \emph{SPARQL}-based way to formulate and check constraints, is generally seen as the method of choice to validate data according to certain constraints \cite{Fuerber2010}, 
although, it is not ideal for their formulation. 
In contrast, high-level constraint languages are comparatively easy to understand and constraints can be formulated more concisely.
Declarative languages may be placed on top of \emph{SPARQL} and \emph{SPIN} when using them as implementation languages.

@@

RDF validation and \emph{OWL 2} assume different semantics. 
This ambiguity in semantics is one of the main reasons why \emph{OWL 2} has not been adopted as a standard constraint language for RDF validation in the past.
We compare semantics of RDF validation and \emph{OWL 2} as \emph{OWL 2} is considered as a constraint language 
when validation semantics is applied.

@@

\emph{OWL 2} is based on the {\em open-world assumption} (\emph{OWA}), i.e., a statement cannot be inferred to be false if it cannot be proved to be true  which fits its primary design purpose to represent knowledge on the \emph{WWW}. 
As each \emph{book} must have a \emph{title} {\small\ms{(Book $\sqsubseteq \exists$ title.$\top$)}} and {\em Hamlet} is a \emph{book}, %{\small\ms{(Book(Hamlet)),}}
{\em Hamlet} must have at least one \emph{title}.
In an \emph{OWA} setting, this constraint does not cause a violation, even if there is no explicitly defined \emph{title}, since there must be a \emph{title} for this \emph{book} which we may not know. 

@@

many RDF validation scenarios require the {\em closed-world assumption} (\emph{CWA}), i.e., a statement is inferred to be false if it cannot be proved to be true.
Thus, classical constraint languages are based on the \emph{CWA} where constraints need to be satisfied only by named individuals. 
The \emph{CWA} yields to a violation since there is no explicitly defined \emph{title} for the \emph{book} {\em Hamlet}. 

@@

\emph{OWL 2} is based on the {\em non-unique name assumption (nUNA)}.  
If we define the property \emph{title} to be functional {\small\ms{(funct (title)),}} a \emph{book} can have at most one distinct \emph{title}.
\emph{UNA} causes a clash
if the \emph{book} {\em Huckleberry-Finn} has more than one \emph{title}.
For \emph{nUNA}, however, reasoning concludes that the \emph{title} {\em The-Adventures-of-Huckleberry-Finn} must be the same as the \emph{title} {\em Die-Abenteuer-des-Huckleberry-Finn} 

@@

%Although, \emph{OWL 2} does not assume \emph{UNA}, \emph{OWL 2} has the constructs \emph{owl:sameAs} and \emph{owl:differentFrom} to state that two names are the same or different.
If validation would assume \emph{OWA} and \emph{nUNA}, validation won’t be that restrictive and therefore we won’t get the intended validation results.

@@

%\emph{OWL 2} reasoning is based on the {\em open-world assumption (OWA)}, i.e., a statement cannot be inferred to be false if it cannot be proved to be true. 

@@

%As each \emph{book} must have a \emph{title} {\small\ms{(Book $\sqsubseteq \exists$ title.$\top$)}} and {\em Hamlet} is a \emph{book} {\small\ms{(Book(Hamlet)),}}
%{\em Hamlet} must have at least one \emph{title}.
%In an \emph{OWA} setting, this axiom does not cause a constraint violation (even if there is no explicitly defined \emph{title}), since there must be a \emph{title} for this \emph{book} which we may not know. 

@@

%When assuming the \emph{CWA}, a constraint violation is raised as there is no explicitly defined title. 
%A constraint that says that a book must have at least one title in \emph{OWA}, 

@@

%But this does not correspond to the intended constraint which is usual in \emph{CWA} way.   
%%Classical constraint languages are based on the \emph{CWA} where constraints need to be satisfied only by named individuals which yields a constraint violation.
%This ambiguity in semantics is one of the main reasons why \emph{OWL 2} has not been adopted as a standard constraint language for RDF validation in the past. 
%In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (\emph{CWA} and \emph{UNA}) as for RDF validation in general. 
%In case we want to use \emph{OWL 2} as a constraint language, we adopt the same semantics (CWA) that is for RDF validation in general. 

@@

as for 56.8\% of the constraint types validation results differ if the \emph{CWA} or the \emph{OWA} 
and as for 66.6\% of the constraint types validation results are different in case the \emph{UNA} or the \emph{nUNA} is assumed \cite{BoschNolleAcarEckert2015}.

@@

%We defined the \textbf{formal generic semantics of the vocabulary} to describe constraints of any constraint type in a generic way (see paper appendix):
%(1) We mapped constraint types to \emph{DL} to logically underpin them and 
%(2) we described the constraint types' intended behavior in terms of validation results depending on \emph{CWA} and \emph{UNA}.
%To be able to define mappings between \emph{specific constraints} (expressed by different languages) and \emph{generic constraints},
%we show for each constraint type that a constraint expressed by a particular language leads to the same answer as the corresponding \emph{DL} query.
%To be able to specify transformations between \emph{specific constraints} (expressed by different languages),
%identical semantics of the languages with regard to the particular constraint type has to be ensured.
%%Constraint languages assume identical semantics, 
%%if for each constraint the following condition holds:
%If a constraint is represented by \emph{ShEx} and \emph{OWL 2}, 
%the validation results must be exactly the same, i.e.,
%whenever the \emph{specific constraint} \emph{$sc_{ShEx}$} is violated, its \emph{$sc_{OWL 2}$} correspondence is also violated which shows that they behave in the same way.
%We expressed constraint types by different languages
%to document identical semantics for these languages with regard to appropriate constraint types.

@@

%\emph{Complex constraints} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Complex constraints} of the constraint type \emph{context-specific exclusive or of property groups} (\emph{R-13}) 
restrict individuals of given classes to have all properties of exactly one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a \emph{title} (for books) or by an \emph{ISSN} and a \emph{title} (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}. 
As \emph{The-Great-Gatsby} is a \emph{publication} with an \emph{ISBN} and a \emph{title} without an \emph{ISSN}, 
\emph{The-Great-Gatsby} is considered as a valid \emph{publication}.
The \emph{complex constraint} is generically expressible by \emph{DL}:
%This \emph{complex constraint} is expressible specifically by \emph{ShEx} and generically by \emph{DL} (and can therefore be mapped to the \emph{RDF-CV} \cite{BoschEckert2015-2}).

@@

In this section, we present current languages for constraint formulation and RDF data validation\footnote{
\emph{SPARQL}: \url{http://www.w3.org/TR/sparql11-query}, 
\emph{SPIN}: \url{http://spinrdf.org},
\emph{Pellet ICV}: \url{http://clarkparsia.com/pellet/icv},
\emph{Stardog ICV}: \url{http://docs.stardog.com/#_validating_constraints},
\emph{ShEx}: \url{http://www.w3.org/Submission/shex-defn},
\emph{ReSh}: \url{http://www.w3.org/Submission/shapes},
\emph{DCMI RDF-AP}: \url{http://dublincore.org/documents/singapore-framework},
\emph{Bibframe}: \url{http://bibframe.org},
\emph{DSP}: \url{http://dublincore.org/documents/dc-dsp},
\emph{RDF Schemarama}: \url{http://swordfish.rdfweb.org/discovery/2001/01/schemarama}
%\emph{Schematron}: \url{http://www.schematron.com},
%\emph{XML Schema}: \url{http://www.w3.org/TR/xmlschema-1},
}.
\emph{SPIN}, \emph{SPARQL}, \emph{OWL 2}, \emph{ShEx}, \emph{ReSh}, and \emph{DSP} are the six most promising and applied languages.
%
The query language for RDF \emph{SPARQL} is generally seen as the method of choice to validate data according to certain constraints \cite{Fuerber2010}, 
although, it is not ideal for their formulation. 
In contrast, high-level constraint languages are comparatively easy to understand and constraints can be formulated more concisely.
Declarative languages may be placed on top of \emph{SPARQL} and \emph{SPIN} when using them as implementation languages.
%
\emph{SPIN} provides a vocabulary to represent \emph{SPARQL} queries as RDF triples
and uses \emph{SPARQL} to specify logical constraints and inference rules \cite{Fuerber2010}.
%
\emph{Stardog ICV} and \emph{Pellet ICV} use \emph{OWL 2} constructs to formulate constraints.
The \emph{Pellet Integrity Constraint Validator (ICV)} is a proof-of-concept extension for the \emph{OWL 2 DL} reasoner \emph{Pellet}. 
\emph{Stardog ICV} 
validates RDF data stored in a Stardog database according to 
constraints which may be written in \emph{SPARQL}, \emph{OWL}, or \emph{SWRL}. 
%
\emph{ShEx} specifies a language whose syntax and semantics are similar to regular expressions.
\emph{ShEx} associate RDF graphs with labeled patterns called \emph{shapes} 
which are used to express formal constraints on the content of RDF graphs.
%
\emph{ReSh} defines its own vocabulary for specifying \emph{shapes} of RDF resources.
Ryman et al. define \emph{shape} as a description of the set of triples a resource is expected to contain and of the integrity constraints those triples are required to satisfy \cite{Ryman2013}. 
%
\emph{DCMI RDF Application Profiles (AP)} and \emph{Bibframe} are approaches to specify profiles for application-specific purposes. 
\emph{DCMI RDF-AP} uses \emph{DSP} as generic constraint language which is also intuitive for non-experts.
The \emph{Bibliographic Framework Initiative (Bibframe)} defines a vocabulary which has a strong overlap with \emph{DSP}.
%
Kontokostas et al. define 17 data quality integrity constraints represented as \emph{SPARQL} query templates called \emph{Data Quality Test Patterns (DQTP)} \cite{Kontokostas2014}. 
%
\emph{RDF Schemarama} 
is a proof-of-concept demonstrator based on ideas from \emph{Schematron}. 
It allows to check that RDF data has required properties and
it is based on \emph{Squish}, an SQL-ish query language for RDF, instead of \emph{SPARQL}. 
%
%\emph{Schematron} is an ISO standard for the validation and quality control of XML documents based on XPath and XSLT. 
%
%\emph{XML Schema} is the primary technology for specifying and constraining the structure of XML documents.
%
In addition to the formulation of constraints, \emph{SPIN} (open source API), \emph{Stardog ICV} (as part of the \emph{Stardog RDF database}), \emph{DQTP} (tests), \emph{Pellet ICV} (extension of \emph{Pellet OWL 2 DL} reasoner) and \emph{ShEx} offer executable validation systems using \emph{SPARQL} as implementation language.

>>>>

In 2013, the \emph{W3C} organized the \emph{RDF Validation Workshop}\footnote{\url{http://www.w3.org/2012/12/rdf-val/}}
where experts from industry, government, and academia discussed first RDF validation use cases. 
%for RDF data validation and constraint formulation.
In 2014, two working groups on RDF validation have been established: 
the \emph{W3C RDF Data Shapes Working Group}\footnote{\url{http://www.w3.org/2014/rds/charter}} and the \emph{DCMI RDF Application Profiles Task Group}\footnote{\url{http://wiki.dublincore.org/index.php/RDF-Application-Profiles}}. 
We collected the findings of these working groups and initiated a database of RDF validation requirements\footnote{Online available at \url{http://purl.org/net/rdf-validation}}
with the intention to collaboratively collect case studies, use cases, requirements, and solutions in a comprehensive and structured way \cite{BoschEckert2014}. 
Based on our work in the \emph{DCMI} and in cooperation with the \emph{W3C} working group,
we published by today 81 requirements to validate RDF data and to formulate constraints; 
%each of them corresponding to a constraint type which in most cases can be expressed by multiple constraint languages. 
each of them corresponds to a constraint type from which concrete constraints are instantiated to be checked on RDF data. 
%and which in most cases can be expressed by multiple constraint languages.
%Requirements/constraint types are uniquely identified by alphanumeric technical identifiers.
%\emph{Minimum qualified cardinality restrictions} which corresponds to the requirement \emph{R-75}, e.g., is a constraint type which corresponds to the requirement {\small\emph{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES}}.
We recently published a technical report (serving as first appendix of this paper) in which we explain each constraint type in detail and give examples for each represented by different constraint languages \cite{BoschNolleAcarEckert2015}.
More than 10 constraint languages with different syntax and semantics exist or are currently developed. 
We focus on the five most promising ones on being the standard like
\emph{Shape Expressions (ShEx)}, \emph{Resource Shapes (ReSh)}, and \emph{Description Set Profiles (DSP)}. 
The \emph{Web Ontology Language} \emph{(OWL)} in its current version OWL 2 can also be used as a constraint language.
With its direct support of validation via \emph{SPARQL}, the \emph{SPARQL Inferencing Notation (SPIN)} is very popular and certainly plays an important role for the future development in this field. 
%\emph{SPIN} is particularly interesting as a means to validate arbitrary constraint languages by mapping them to \emph{SPARQL} \cite{BoschEckert2014-2}.

@@

\begin{table}[H]
	\centering
	  %\caption{Constraint Type Specific Expressivity}
		\caption{Constraint Type Specific Expressivity of Constraint Languages}
	  \scriptsize
		\begin{tabular}{c|c|c|c|c}
      %\textbf{Expressivity Class} & $Expr(\overline{\mathcal{CT}_R})$ (42) & $Expr(\mathcal{CT}_R)$ (32) \\		
			\emph{DSP} & \emph{ReSh} & \emph{ShEx} & OWL 2 & \emph{SPIN} \\	
      \hline
			17.3 (14) & 25.9 (21) & 29.6 (24) & 67.9 (55) & \textbf{100.0} \textbf{(81)} 
		\end{tabular}
	\label{tab:constraint-type-specific-expressivity}
\end{table}

The fact that \emph{SPIN} supports all and OWL 2 covers 2/3 of all constraint types
emphasizes the significant role \emph{SPIN} and OWL 2 play for the future development of constraint languages.
\emph{SPARQL} and therefore \emph{SPIN}, the \emph{SPARQL}-based way to formulate and check constraints, is generally seen as the method of choice to validate data according to certain constraints \cite{Fuerber2010}, 
although, it is not ideal for their formulation. 
In contrast, high-level constraint languages are comparatively easy to understand and constraints can be formulated more concisely.
Declarative languages may be placed on top of \emph{SPARQL} and \emph{SPIN} when using them as implementation languages.

@@

RDF validation and OWL 2 assume different semantics. 
This ambiguity in semantics is one of the main reasons why OWL 2 has not been adopted as a standard constraint language for RDF validation in the past.
We compare semantics of RDF validation and OWL 2 as OWL 2 is considered as a constraint language 
in case the semantics for RDF validation is applied.

@@

OWL 2 is based on the {\em open-world assumption} (OWA), i.e., a statement cannot be inferred to be false if it cannot be proved to be true  which fits its primary design purpose to represent knowledge on the World Wide Web. 
As each book must have a title {\small\ms{(Book $\sqsubseteq \exists$ title.$\top$)}} and {\em Hamlet} is a book, %{\small\ms{(Book(Hamlet)),}}
{\em Hamlet} must have a title as well.
In an OWA setting, this constraint does not cause a violation, even if there is no explicitly defined title, since there must be a title for this book which we may not know. 

@@

many RDF validation scenarios require the {\em closed-world assumption} (CWA), i.e., a statement is inferred to be false if it cannot be proved to be true.
Thus, classical constraint languages are based on the CWA where constraints need to be satisfied only by named individuals. 
The CWA yields to a violation since there is no explicitly defined title for the book {\em Hamlet}. 

@@

OWL 2 is based on the {\em non-unique name assumption (nUNA)}.  
If we define the property title to be functional {\small\ms{(funct (title)),}} a book can have at most one distinct title.
UNA causes a clash
if the book {\em Huckleberry-Finn} has more than one title.
For nUNA, however, reasoning concludes that the title {\em The-Adventures-of-Huckleberry-Finn} must be the same as the title {\em Die-Abenteuer-des-Huckleberry-Finn} 

@@

%Although, OWL 2 does not assume UNA, OWL 2 has the constructs \emph{owl:sameAs} and \emph{owl:differentFrom} to state that two names are the same or different.
If validation would assume OWA and nUNA, validation won’t be that restrictive and therefore we won’t get the intended validation results.

@@

%OWL 2 reasoning is based on the {\em open-world assumption (OWA)}, i.e., a statement cannot be inferred to be false if it cannot be proved to be true. 

@@

%As each book must have a title {\small\ms{(Book $\sqsubseteq \exists$ title.$\top$)}} and {\em Hamlet} is a book {\small\ms{(Book(Hamlet)),}}
%{\em Hamlet} must have at least one title.
%In an OWA setting, this axiom does not cause a constraint violation (even if there is no explicitly defined title), since there must be a title for this book which we may not know. 

@@

%When assuming the CWA, a constraint violation is raised as there is no explicitly defined title. 
%A constraint that says that a book must have at least one title in OWA, 

@@

%But this does not correspond to the intended constraint which is usual in CWA way.   
%%Classical constraint languages are based on the CWA where constraints need to be satisfied only by named individuals which yields a constraint violation.
%This ambiguity in semantics is one of the main reasons why OWL 2 has not been adopted as a standard constraint language for RDF validation in the past. 
%In case we want to use OWL 2 as a constraint language, we adopt the same semantics (CWA and UNA) as for RDF validation in general. 

@@

as for 56.8\% of the constraint types validation results differ if the CWA or the OWA 
and as for 66.6\% of the constraint types validation results are different in case the UNA or the nUNA is assumed \cite{BoschNolleAcarEckert2015}.

@@

%We defined the \textbf{formal generic semantics of the vocabulary} to describe constraints of any constraint type in a generic way (see paper appendix):
%(1) We mapped constraint types to \emph{DL} to logically underpin them and 
%(2) we described the constraint types' intended behavior in terms of validation results depending on CWA and UNA.
%To be able to define mappings between \emph{specific constraints} (expressed by different languages) and \emph{generic constraints},
%we show for each constraint type that a constraint expressed by a particular language leads to the same answer as the corresponding \emph{DL} query.
%To be able to specify transformations between \emph{specific constraints} (expressed by different languages),
%identical semantics of the languages with regard to the particular constraint type has to be ensured.
%%Constraint languages assume identical semantics, 
%%if for each constraint the following condition holds:
%If a constraint is represented by \emph{ShEx} and OWL 2, 
%the validation results must be exactly the same, i.e.,
%whenever the \emph{specific constraint} \emph{$sc_{ShEx}$} is violated, its \emph{$sc_{OWL 2}$} correspondence is also violated which shows that they behave in the same way.
%We expressed constraint types by different languages
%to document identical semantics for these languages with regard to appropriate constraint types.

@@

%\emph{Complex constraints} are composed out of \emph{simple constraints} and/or \emph{complex constraints}.
\emph{Complex constraints} of the constraint type \emph{context-specific exclusive or of property groups} (\emph{R-13}) 
restrict individuals of given classes to have all properties of exactly one of multiple mutually exclusive property groups.
\emph{Publications}, e.g., are either identified by an \emph{ISBN} and a title (for books) or by an \emph{ISSN} and a title (for periodical \emph{publications}), but it should not be possible to assign both identifiers to a given \emph{publication}. 
As \emph{The-Great-Gatsby} is a \emph{publication} with an \emph{ISBN} and a title without an \emph{ISSN}, 
\emph{The-Great-Gatsby} is considered as a valid \emph{publication}.
The \emph{complex constraint} is generically expressible by \emph{DL}:
%This \emph{complex constraint} is expressible specifically by \emph{ShEx} and generically by \emph{DL} (and can therefore be mapped to the \emph{RDF-CV} \cite{BoschEckert2015-2}).

@@

In this section, we present current languages for constraint formulation and RDF data validation\footnote{
\emph{SPARQL}: \url{http://www.w3.org/TR/sparql11-query}, 
\emph{SPIN}: \url{http://spinrdf.org},
\emph{Pellet ICV}: \url{http://clarkparsia.com/pellet/icv},
\emph{Stardog ICV}: \url{http://docs.stardog.com/#_validating_constraints},
\emph{ShEx}: \url{http://www.w3.org/Submission/shex-defn},
\emph{ReSh}: \url{http://www.w3.org/Submission/shapes},
\emph{DCMI RDF-AP}: \url{http://dublincore.org/documents/singapore-framework},
\emph{Bibframe}: \url{http://bibframe.org},
\emph{DSP}: \url{http://dublincore.org/documents/dc-dsp},
\emph{RDF Schemarama}: \url{http://swordfish.rdfweb.org/discovery/2001/01/schemarama}
%\emph{Schematron}: \url{http://www.schematron.com},
%\emph{XML Schema}: \url{http://www.w3.org/TR/xmlschema-1},
}.
\emph{SPIN}, \emph{SPARQL}, OWL 2, \emph{ShEx}, \emph{ReSh}, and \emph{DSP} are the six most promising and applied languages.
%
The query language for RDF \emph{SPARQL} is generally seen as the method of choice to validate data according to certain constraints \cite{Fuerber2010}, 
although, it is not ideal for their formulation. 
In contrast, high-level constraint languages are comparatively easy to understand and constraints can be formulated more concisely.
Declarative languages may be placed on top of \emph{SPARQL} and \emph{SPIN} when using them as implementation languages.
%
\emph{SPIN} provides a vocabulary to represent \emph{SPARQL} queries as RDF triples
and uses \emph{SPARQL} to specify logical constraints and inference rules \cite{Fuerber2010}.
%
\emph{Stardog ICV} and \emph{Pellet ICV} use OWL 2 constructs to formulate constraints.
The \emph{Pellet Integrity Constraint Validator (ICV)} is a proof-of-concept extension for the \emph{OWL 2 DL} reasoner \emph{Pellet}. 
\emph{Stardog ICV} 
validates RDF data stored in a Stardog database according to 
constraints which may be written in \emph{SPARQL}, \emph{OWL}, or \emph{SWRL}. 
%
\emph{ShEx} specifies a language whose syntax and semantics are similar to regular expressions.
\emph{ShEx} associate RDF graphs with labeled patterns called \emph{shapes} 
which are used to express formal constraints on the content of RDF graphs.
%
\emph{ReSh} defines its own vocabulary for specifying \emph{shapes} of RDF resources.
Ryman et al. define \emph{shape} as a description of the set of triples a resource is expected to contain and of the integrity constraints those triples are required to satisfy \cite{Ryman2013}. 
%
\emph{DCMI RDF Application Profiles (AP)} and \emph{Bibframe} are approaches to specify profiles for application-specific purposes. 
\emph{DCMI RDF-AP} uses \emph{DSP} as generic constraint language which is also intuitive for non-experts.
The \emph{Bibliographic Framework Initiative (Bibframe)} defines a vocabulary which has a strong overlap with \emph{DSP}.
%
Kontokostas et al. define 17 data quality integrity constraints represented as \emph{SPARQL} query templates called \emph{Data Quality Test Patterns (DQTP)} \cite{Kontokostas2014}. 
%
\emph{RDF Schemarama} 
is a proof-of-concept demonstrator based on ideas from \emph{Schematron}. 
It allows to check that RDF data has required properties and
it is based on \emph{Squish}, an SQL-ish query language for RDF, instead of \emph{SPARQL}. 
%
%\emph{Schematron} is an ISO standard for the validation and quality control of XML documents based on XPath and XSLT. 
%
%\emph{XML Schema} is the primary technology for specifying and constraining the structure of XML documents.
%
In addition to the formulation of constraints, \emph{SPIN} (open source API), \emph{Stardog ICV} (as part of the \emph{Stardog RDF database}), \emph{DQTP} (tests), \emph{Pellet ICV} (extension of \emph{Pellet OWL 2 DL} reasoner) and \emph{ShEx} offer executable validation systems using \emph{SPARQL} as implementation language.
